---
title: "DataCleaningExample"
author: "Sousa"
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
library(dplyr); library(tidyr)
```

# Read in data

We are using data from: 
[Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2. ](https://knb.ecoinformatics.org/view/df35b.304.2)



```{r}
catch_original = read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1"), stringsAsFactors = F)

head(catch_original)
```

# Clean and reshape
##Clean
Remove marginal sum and notes columns using `select()`
```{r}
catch_clean = catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% 
  #same as select(-All, -notesRegCode) 
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))

head(catch_clean)
```
## QA

Find where as.numeric couldn't convert Chinook to a number
```{r}
i = which(is.na(catch_clean$Chinook))
i
```


##Reshape
Reshape from Wide to Tall format using `pivot_longer`

```{r}
catch_long = catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "species",
               values_to = "catch") %>% 
  rename(catch_thousands = catch) %>% 
  mutate(catch = catch_thousands * 1000) %>% 
  select(-catch_thousands)

head(catch_long)
```

example with `pivot_wider`

```{r}
catch_wide = catch_long %>% 
  pivot_wider(names_from = Year,
              values_from = catch)

head(catch_wide)
```

## group_by and summarize
```{r}
mean_region = catch_long %>% 
  group_by(Region, species) %>% 
  summarise(catch_mean = mean(catch),
            num_obs = n())

head(mean_region)
```

```{r}
mean_species = catch_long %>% 
  group_by(species) %>% 
  summarise(mean_catch = mean(catch),
            num_obs = n())
head(mean_species)
```

What year had the highest salmon catch?

```{r}
annual_catch = catch_long %>% 
  group_by(Year) %>% 
  summarise(total_catch = sum(catch)) %>% 
  arrange(-total_catch)

annual_catch
```

What year had the highest sockeye catch?


```{r}
annual_sockeye = catch_long %>% 
  filter(species == "Sockeye") %>% 
  group_by(Year) %>% 
  summarise(total_catch = sum(catch)) %>% 
  arrange(-total_catch)

head(annual_sockeye)
```


# Join to regions table
```{r}

```



Read in regions data

```{r}
region_defs = read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",stringsAsFactors = F) %>% 
  select(code, mgmtArea)

head(region_defs)
```

```{r}
catch_joined = left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)

```

