---
title: "SpatialDataExample"
author: "Sousa"
date: "2/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
# First load libraries
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(leaflet)
library(scales)
library(ggmap)
```


# Read in and inspect data
```{r}
## Read in shapefile using sf
ak_regions <- read_sf("data/ak_regions_simp.shp")

plot(ak_regions)

class(ak_regions)
head(ak_regions)
st_crs(ak_regions)
```

## Transform CRS
```{r}
#Use st_transform to transfrom from WGS84 to Alaska Albers
ak_regions_3338 = ak_regions %>% 
  st_transform(crs = 3338)

plot(ak_regions_3338)
```

## Works with TidyVerse
```{r}
ak_regions_3338 %>% 
  filter(region == "Southeast") %>% 
  select(region)
```
# Spatial joins and summaries

```{r}
#Read in population data
pop = read.csv("data/alaska_population.csv", stringsAsFactors = F)

head(pop)
#it's not yet a spatial object - has no projection

#so set as sf object and give CRS
pop_4326 = st_as_sf(pop,
                    coords = c('lng', 'lat'),
                    crs = 4326,
                    remove = F)

head(pop_4326)

pop_3338 = pop_4326 %>% 
  st_transform(crs = 3338)
```

Now do spatial join to figure out which city is in which region
```{r}
pop_joined = st_join(pop_3338, ak_regions_3338, join = st_within)
head(pop_joined)
plot(pop_joined)
```

## calculate population by region
```{r}
#you need to set as data frame first to get rid of geometry. otherwise you end up with a long list of points that are each assigned the value of the sum of the population in that region, rather than just one value
pop_region = pop_joined %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarize(total_pop = sum(population))

head(pop_region)
```

```{r}
#now do a left join to stick on the polygons so you can make the plot
pop_region_3338 = left_join(ak_regions_3338, pop_region, by = "region")
plot(pop_region_3338)
```

# Make mapzzz

```{r}
# now just need to make the pretty plot
pop_mgmt_3338 = pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarize(total_pop = sum(total_pop), do_union = T) 

plot(pop_mgmt_3338["total_pop"])
```

## Write out objects
```{r}
#delete_layer = TRUE means overwrite whole file with this output, rather than add the result as a new layer
write_sf(pop_region_3338, "data/ak_regions_pop_output.shp", delete_layer = TRUE)
```

## Extra pretty plot
```{r}
rivers_3338 = read_sf("data/ak_rivers_simp.shp")
st_crs(rivers_3338)
```


```{r}
#labels = comma gets rid of the scientific notation
ggplot() + 
  geom_sf(data = pop_region_3338, aes(fill = total_pop)) +
  geom_sf(data = pop_3338, aes(), size = 0.5) +
  geom_sf(data = rivers_3338, aes(size = StrOrder), color = "black") +
  scale_size(range = c(0.01, 0.2), guide = F) +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high = "firebrick", labels = comma)
    
```


# Getting basemaps using ggmap

## First transform pop into new CRS.
```{r}
pop_3857 = pop_3338 %>% 
  st_transform(crs = 3857)


```

## Function to fix weird problem with Stamen Maps
```{r}
# Define a function to fix the bbox to be in EPSG:3857
# See https://github.com/dkahle/ggmap/issues/160#issuecomment-397055208
ggmap_bbox_to_3857 <- function(map) {
  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
  # and set the names to what sf::st_bbox expects:
  map_bbox <- setNames(unlist(attr(map, "bb")), 
                       c("ymin", "xmin", "ymax", "xmax"))
  
  # Coonvert the bbox to an sf polygon, transform it to 3857, 
  # and convert back to a bbox (convoluted, but it works)
  bbox_3857 <- st_bbox(st_transform(st_as_sfc(st_bbox(map_bbox, crs = 4326)), 3857))
  
  # Overwrite the bbox of the ggmap object with the transformed coordinates 
  attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
  attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
  attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
  attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
  map
}
```

## Get Stamen Map
```{r}
bbox = c(-170,52,-130,64)
ak_map = get_stamenmap(bbox, zoom = 4)
ak_map_3857 = ggmap_bbox_to_3857(ak_map)
```

## Plot Stamen Map
```{r}
ggmap(ak_map_3857) +
  geom_sf(data = pop_3857, aes(color = population), inherit.aes = F) +
  scale_color_continuous(low="khaki", high="firebrick", labels = comma, trans = "log10")
```

## Interactive mapz using leaflet
```{r}
#define leaflet projection 
epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(16:7))
```

```{r}
#need to transform pop CRS again
pop_region_4326 = pop_region_3338  %>% 
  st_transform(crs = 4326)
```

```{r}
pal <- colorNumeric(palette = "Reds", domain = pop_region_4326$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
        addPolygons(data = pop_region_4326, 
                    fillColor = ~pal(total_pop),
                    weight = 1,
                    color = "black",
                    fillOpacity = 1) %>% 
        addCircleMarkers(data = pop_4326,
                         lat = ~lat,
                         lng = ~lng,
                         radius = ~log(population/500), # arbitrary scaling
                         fillColor = "gray",
                         fillOpacity = 1,
                         weight = 0.25,
                         color = "black",
                         label = ~paste0(pop_4326$city, ", population ", comma(pop_4326$population))) %>%
        addLegend(position = "bottomleft",
                  pal = pal,
                  values = range(pop_region_4326$total_pop),
                  title = "Total Population")

m
```
